Pattern matching for\\ $(213,231)$-avoiding permutations

### 2015-02-09 article.tex
end cases

\begin{abstract}
	Given permutations $\sigma \in S_k$ and $\pi \in S_n$, the Permutation Pattern (PP)
	problem is to decide whether $\pi$ contains $\sigma$ as an order-isomorphic pattern.
	We give a linear-time algorithm in case both $\pi$ and $\sigma$ avoid
	$213$ and $231$.
	For the special where only $\sigma$ avoids $213$ and $231$, we present a
	$O(kn^2)$ time algorithm. We extend our research to $(213,231)$-avoiding
	bivincular pattern and present a $O(kn^3)$ time algorithm.
\end{abstract}

@@

	A permutation $\pi$ is said to contain another permutation $\sigma$,
	in symbols $\sigma \preceq \pi$,
	if there exists a subsequence of entries of $\pi$ that has the same relative
	order as $\sigma$, and in this case $\sigma$ is said to be a
	\emph{pattern} of $\pi$.
	Otherwise, $\pi$ is said to avoid the permutation $\sigma$.
	For example a permutation contains the pattern $123$ (resp. $321$) if it has
	an increasing (resp. decreasing) subsequence of length $3$.
	Here, note that members need not actually to be consecutive,
	merely increasing (resp. descreasing).
	As another example,
	$6152347$ contains $213$ but not $231$.
	During the last decade, the study of the pattern containment on permutations has
	become a very active area of research and
	a whole annual conference (\textsc{Permutation Patterns}) is now devoted
	to this topic.

@@

	A few particular cases of the pattern containment problem have been attacked successfully.
	The case of increasing patterns is solvable in
	$O(|\pi| \log \log |\sigma|)$ time in the RAM model \cite{Crochemore:Porat:2010},
	improving the previous 30-year bound of $O(|\pi| \log |\sigma|)$.
	(The algorithm also improves on the previous
	$O(|\pi| \log \log |\pi|)$ bound.)
	Furthermore, the patterns $132$, $213$, $231$, $312$ can all be handled in linear-time
	by stack sorting algorithms.
	Any pattern of length $4$ can be detected in $O(|\pi| \log |\pi|)$ time
	\cite{Albert:Aldred:Atkinson:Holton:ISAAC:2001}.
	Algorithmic issues for $321$-avoiding patterns containment has been investigated in
	\cite{Guillemot:Vialette:ISAAC:2009}.
	The pattern containment problem is also solvable in
	polynomial-time for separable patterns \cite{Ibarra:1997,Bose:Buss:Lubiw:1998}
	(see also \cite{Bouvel:Rossin:Vialette:CPM:2007} for LCS-like issues
	of separable permutations).
	Separable permutations are those permutations that contain neither
	$2413$ nor $3142$, and they are enumerated by the Schröder numbers
	(sequence A006318 in OEIS).
	To the best of our knowledge,
	separable permutations first arose in the work of
	Avis and Newborn~\cite{Avis:Newborn:1981},
	who showed that they are precisely the permutations which can be sorted by an
	arbitrary number of pop-stacks in series,
	where a pop-stack is a restricted form of stack in which any pop operation
	pops all items at once.
	(Notice that the separable permutations include as a special case the
	stack-sortable permutations, which avoid the pattern $231$.)

@@

A \emph{permutation} of length $n$ is a one-to-one function from an
$n$-element set to itself.
We write permutations as words
$\sigma = \sigma_1\sigma_2\,\ldots\,\sigma_n$, whose letters are distinct
and uasually consist of the integers $12\,\ldots\,n$.
We let
$\pi[i]$ stands for $\pi_i$.
For the sake of convenience, we let
$\pi[i:j]$ stand for
$\pi_i\pi_{i+1}\,\ldots\,\pi_j$,
$\pi[:j]$ stand for $\pi[1:j]$ and
$\pi[i:]$ stand for $\pi[i:n]$.
As usual, we let $S_n$ denote the set of all permutations of length $n$.

>>>>

\begin{abstract}
	Given permutations $\sigma \in S_k$ and $\pi \in S_n$, the
  \emph{pattern containment} problem is to decide whether $\pi$ contains
  $\sigma$ as an order-isomorphic subsequence.
	We give a linear-time algorithm in case both $\pi$ and $\sigma$ avoid
	$213$ and $231$.
	For the special where only $\sigma$ avoids $213$ and $231$, we present a
	$O(kn^2)$ time algorithm. We extend our research to $(213,231)$-avoiding
	bivincular pattern and present a $O(kn^3)$ time algorithm.
\end{abstract}

@@

	A permutation $\pi$ is said to contain another permutation $\sigma$,
	in symbols $\sigma \preceq \pi$,
	if there exists a subsequence of entries of $\pi$ that has the same relative
	order as $\sigma$, and in this case $\sigma$ is said to be a
	\emph{pattern} of $\pi$.
	Otherwise, $\pi$ is said to \emph{avoid} the permutation $\sigma$.
	For example a permutation contains the pattern $123$ (resp. $321$) if it has
	an increasing (resp. decreasing) subsequence of length $3$.
	Here, note that members need not actually to be consecutive,
	merely increasing (resp. descreasing).
	As another example,
	$6152347$ contains $213$ but not $231$.
	During the last decade, the study of the pattern containment on permutations has
	become a very active area of research and
	a whole annual conference (\textsc{Permutation Patterns}) is now devoted
	to this topic.

@@

	A few particular cases of the pattern containment problem have been attacked successfully.
	The case of increasing patterns is solvable in
	$O(|\pi| \log \log |\sigma|)$ time in the RAM model \cite{Crochemore:Porat:2010},
	improving the previous 30-year bound of $O(|\pi| \log |\sigma|)$.
	(The algorithm also improves on the previous
	$O(|\pi| \log \log |\pi|)$ bound.)
	Furthermore, the patterns $132$, $213$, $231$, $312$ can all be handled in linear-time
	by stack sorting algorithms.
	Any pattern of length $4$ can be detected in $O(|\pi| \log |\pi|)$ time
	\cite{Albert:Aldred:Atkinson:Holton:ISAAC:2001}.
	Algorithmic issues for $321$-avoiding patterns containment has been investigated in
	\cite{Guillemot:Vialette:ISAAC:2009}.
	The pattern containment problem is also solvable in
	polynomial-time for separable patterns \cite{Ibarra:1997,Bose:Buss:Lubiw:1998}
	(see also \cite{Bouvel:Rossin:Vialette:CPM:2007} for LCS-like issues
	of separable permutations).
	Separable permutations are those permutations that contain neither
	$2413$ nor $3142$, and they are enumerated by the Schr{\"o}der numbers
	(sequence A006318 in OEIS).
	To the best of our knowledge,
	separable permutations first arose in the work of
	Avis and Newborn~\cite{Avis:Newborn:1981},
	who showed that they are precisely the permutations which can be sorted by an
	arbitrary number of pop-stacks in series,
	where a pop-stack is a restricted form of stack in which any pop operation
	pops all items at once.
	(Notice that the separable permutations include as a special case the
	stack-sortable permutations, which avoid the pattern $231$.)

@@

A \emph{permutation} of length $n$ is a one-to-one function from an
$n$-element set to itself.
We write permutations as words
$\pi = \pi_1\pi_2\,\ldots\,\pi_n$, whose letters are distinct
and uasually consist of the integers $12\,\ldots\,n$.
We let
$\pi[i]$ stands for $\pi_i$.
For the sake of convenience, we let
$\pi[i:j]$ stand for
$\pi_i\pi_{i+1}\,\ldots\,\pi_j$,
$\pi[:j]$ stand for $\pi[1:j]$ and
$\pi[i:]$ stand for $\pi[i:n]$.
As usual, we let $S_n$ denote the set of all permutations of length $n$.
